# Part 1. Prepare Training Data
In 'gibbonR' there are two ways that you can format your training data. The first can be a set of labelled .wav clips with the class indicated in the name of the file (e.g., 'gibbon_01.wav' and 'noise_01.wav'). The second is to have a folder of selection tables created in Raven Pro (K. Lisa Yang Center for Conservation Bioacoustics) and a folder with the associated '.wav' files. For the second approach there must be an annotation column indicating the call type and it is assumed that all signals of interest are annotated, and the rest of the files contain only background noise.  

## Download the data from Github
```{eval=FALSE}
# You need to tell R where to store the zip files on your computer.
destination.file.path.zip <-
  "/Users/denaclink/Downloads/BorneoExampleData.zip"

# You also need to tell R where to save the unzipped files
destination.file.path <- "/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/"

# This function will download the data from github

utils::download.file("https://github.com/DenaJGibbon/BorneoExampleData/archive/master.zip",
                     destfile = destination.file.path.zip)

# This function will unzip the file
utils::unzip(zipfile = destination.file.path.zip,
             exdir = destination.file.path)

# Examine the contents
list.of.sound.files <- list.files(paste(destination.file.path,
                                        "BorneoExampleData-master", "data", sep =
                                          "/"),
                                  full.names = T)
list.of.sound.files

```

Use this function to read in the .RDA file and save it as an R object from https://stackoverflow.com/questions/5577221/how-can-i-load-an-object-into-a-variable-name-that-i-specify-from-an-r-data-file

```{r eval=FALSE}
loadRData <- function(fileName) {
  #loads an RData file, and returns it
  load(fileName)
  get(ls()[ls() != "fileName"])
}
```

This function will load the entire list of r data files
```{r eval=FALSE}
list.rda.files <- list()
for(x in 1:length(list.of.sound.files)){
  list.rda.files[[x]] <-  loadRData(list.of.sound.files[[x]])
}
```

Assign each rda an informative name
```{r eval=FALSE, warning=FALSE}
multi.class.list <- list.rda.files[[1]]
S11_20180219_060002_1800sto3600s <- list.rda.files[[2]]
```

Now we create a directory with the training .wav files
```{r eval=FALSE, warning=FALSE}
TrainingDataDirectory <- "/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/BorneoMultiClass"

for(a in 1:length(multi.class.list)){
  Temp.element <- multi.class.list[[a]]
  writeWave(Temp.element[[2]], paste(TrainingDataDirectory,Temp.element[[1]],sep='/'))
}

```

## Part 1A. Training Data with Labeled .wav clips
### Read in clips and calculate MFCCs
```{r echo = T, results = 'hide' }
TrainingWavFilesDir <- 
  "/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/BorneoMultiClass/"

trainingdata <- gibbonR::MFCCFunction(input.dir=TrainingWavFilesDir, min.freq = 400, max.freq = 1600,win.avg="standard")


trainingdata$class <- as.factor(trainingdata$class)
```

### Compare Random Forest and Support Vector Machine for Supervised Classification
```{r }

trainingdata$class <- as.factor(trainingdata$class)


ml.model.svm <- e1071::svm(trainingdata[, 2:ncol(trainingdata)], trainingdata$class, kernel = "radial", 
                           cross = 25,
                           probability = TRUE)

print(paste('SVM accuracy',ml.model.svm$tot.accuracy))


ml.model.rf <- randomForest::randomForest(x=trainingdata[, 2:ncol(trainingdata)], y = trainingdata$class)


print(ml.model.rf)
```

## Part 1B. Training Data with Raven Selection Tables
### Prepare training data from labeled annotations
```{r,eval=F }
# Specify the folder where the training data will be saved
TrainingDataFolderLocation <- "/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/TrainingDataFromRavenSelectionTables"

# Directory with annotated selection tables
AnnotatedSelectionTables <- list.files("/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/SelectionTables/GibbonTrainingSelectionTables/",
                                       full.names = T)

# Directory with corresponding .wav files
AnnotatedWaveFiles <- list.files("/Users/denaclink/Library/CloudStorage/Box-Box/gibbonRSampleFiles/GibbonTrainingFiles/",full.names = T)
AnnotatedWaveFilesShort <- list.files("/Users/denaclink/Library/CloudStorage/Box-Box/gibbonRSampleFiles/GibbonTrainingFiles/",full.names = F)
AnnotatedWaveFilesShort <- str_split_fixed(AnnotatedWaveFilesShort,pattern = '.wav', n=2)[,1]

# Loop to cut out the corresponding annotations into short clips
for(i in 1: length(AnnotatedSelectionTables)){
  
  # Read in selection table
  TempSelectionTable <- read.delim2(AnnotatedSelectionTables[i])
  
  # Find the corresponding soundfile
  SoundFileIndex <- which(str_detect(AnnotatedSelectionTables[i],AnnotatedWaveFilesShort))
  
  TempAnnotateWave <- readWave(AnnotatedWaveFiles[SoundFileIndex])
  
  ShortSoundClips <- lapply(1:nrow(TempSelectionTable),
                                function(j) extractWave(TempAnnotateWave,
                                                        from= as.numeric(TempSelectionTable[j,]$Begin.Time..s.),
                                                        to=as.numeric(TempSelectionTable[j,]$ End.Time..s.),
                                                        xunit = c("time"),plot=F,output="Wave"))
  # Write wave files to folder
  for(k in 1:length(ShortSoundClips)){
    TempClip <- ShortSoundClips[[k]]
    WavFileName <- paste(TrainingDataFolderLocation,'/female.gibbon_', k, '.wav',sep="")
    writeWave(TempClip,WavFileName,extensible = F)
  }
  
  
}

```

### Prepare noise training data from files without target signal
```{r eval=F}
# Specify the folder where the training data will be saved
TrainingDataFolderLocation <- "/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/TrainingDataFromRavenSelectionTables/"

# Directory with annotated selection tables
NoiseSelectionTables <- list.files("/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/SelectionTables/NoiseSelectionTables/",
                                       full.names = T)

# Directory with corresponding .wav files
NoiseWaveFiles <- list.files("/Users/denaclink/Library/CloudStorage/Box-Box/gibbonRSampleFiles/NoiseFiles/",full.names = T)
NoiseWaveFilesShort <- list.files("/Users/denaclink/Library/CloudStorage/Box-Box/gibbonRSampleFiles/NoiseFiles/",full.names = F)
NoiseWaveFilesShort <- str_split_fixed(NoiseWaveFilesShort,pattern = '.wav', n=2)[,1]

for(i in 1:length(NoiseSelectionTables)){
  
  # Find the corresponding soundfile
  SoundFileIndex <- which(str_detect(NoiseSelectionTables[i],NoiseWaveFilesShort))

  DetectBLED(input=NoiseWaveFiles[SoundFileIndex],
           min.freq = 400, 
           max.freq = 1600,
           noise.quantile.val=0.3,
           spectrogram.window =512,
           pattern.split = ".wav", 
           min.signal.dur = 3,
           max.sound.event.dur = 12, 
           wav.output = "TRUE", 
           output.dir = TrainingDataFolderLocation,
           swift.time=TRUE,
           time.start=06,
           time.stop=11,
           write.table.output=TRUE,
           verbose=TRUE,
           random.sample=FALSE)
}

```

### Now read in clips based on Raven Selection tables and calculate MFCCs
```{r echo = T, results = 'hide' }

TrainingWavFilesDir <- 
  "/Users/denaclink/Desktop/RStudio Projects/gibbonR/data/TrainingDataFromRavenSelectionTables/"

trainingdata <- gibbonR::MFCCFunction(input.dir=TrainingWavFilesDir, min.freq = 400, max.freq = 1600,win.avg="standard")


trainingdata$class <- as.factor(trainingdata$class)
```

### Compare Random Forest and Support Vector Machine for Supervised Classification
```{r }

trainingdata$class <- as.factor(trainingdata$class)


ml.model.svm <- e1071::svm(trainingdata[, 2:ncol(trainingdata)], trainingdata$class, kernel = "radial", 
                           cross = 25,
                           probability = TRUE)

print(paste('SVM accuracy',ml.model.svm$tot.accuracy))


ml.model.rf <- randomForest::randomForest(x=trainingdata[, 2:ncol(trainingdata)], y = trainingdata$class)


print(ml.model.rf)


```
