# Part 3. Calculate performance metrics
## Part 3a. Load libraries and specify file paths
Please note this script is very much in development!
```{r ,warning=FALSE, message=FALSE}
# Load necessary libraries
library(stringr)  # For string manipulation
library(caret)    # For machine learning and model evaluation
library(ggpubr)   # For data visualization
library(dplyr)    # For data manipulation
library(data.table) # For sorting the detections
library(ggplot2)  # For data visualization


# Change this so that it points to the location on your computer
FolderPath <- '/Users/denaclink/Library/CloudStorage/Box-Box/gibbonRSampleFiles/'


# Get a list of TopModel result files
TopModelresults <- list.files('data/DetectAndClassifyOutput',
                              full.names = TRUE)

# Get a list of annotation selection table files
TestDataSet <- list.files( paste(FolderPath,"/SelectionTables/GibbonTestSelectionTables", sep=''),
                          full.names = TRUE)

```

## Part 3b. Match detections with human annotations
Now we match the detections from the model to human annotations. The start time/end time buffer tells us how much flexibility the detections can have from the start/end time of the call. If the calls are short, then this should be short.
```{r, warning=FALSE}
start.time.buffer <- 8
end.time.buffer <- 4

# Preallocate space for TopModelDetectionDF
TopModelDetectionDF <- data.frame()

# Loop through each TopModel result file
for (f in 1:length(TopModelresults)) {

  # Read the TopModel result table into a data frame
  TempTopModelTable <- read.delim2(TopModelresults[f])

  # Extract the short name of the TopModel result file
  ShortName <- basename(TopModelresults[f])
  ShortName <- str_split_fixed(ShortName, pattern = 'gibbonR', n = 2)[, 1]

  # Find the corresponding annotation selection table
  testDataIndex <- which(str_detect(TestDataSet, ShortName))

  if(length(testDataIndex) > 0 ){
  TestDataTable <- read.delim2(TestDataSet[testDataIndex])


  # Round Begin.Time..s. and End.Time..s. columns to numeric
  TestDataTable$Begin.Time..s. <- round(as.numeric(TestDataTable$Begin.Time..s.))
  TestDataTable$End.Time..s. <- round(as.numeric(TestDataTable$End.Time..s.))

  DetectionList <- list()
  # Loop through each row in TempTopModelTable
  for (c in 1:nrow(TempTopModelTable)) {
    TempRow <- TempTopModelTable[c,]

    # Check if Begin.Time..s. is not NA
    if (!is.na(TempRow$Begin.Time..s.)) {
      # Convert Begin.Time..s. and End.Time..s. to numeric
      TempRow$Begin.Time..s. <- as.numeric(TempRow$Begin.Time..s.)
      TempRow$End.Time..s. <- as.numeric(TempRow$End.Time..s.)

      # Determine if the time of the detection is within the time range of an annotation
      TimeBetween <- data.table::between(TempRow$Begin.Time..s.,
                                         TestDataTable$Begin.Time..s. - start.time.buffer,
                                         TestDataTable$End.Time..s. + end.time.buffer)

      # Extract the detections matching the time range
      matched_detections <- TestDataTable[TimeBetween, ]

      if (nrow(matched_detections) > 0) {
        # Set signal based on the Call.Type in matched_detections
        TempRow$signal <- 'female.gibbon'
        DetectionList[[length( unlist(DetectionList))+1]] <-  which(TimeBetween == TRUE)
      } else {
        # Set signal to 'Noise' if no corresponding annotation is found
        TempRow$signal <- 'noise'
      }

      # Append TempRow to TopModelDetectionDF
      TopModelDetectionDF <- rbind.data.frame(TopModelDetectionDF, TempRow)
    }
  }

  # Identify missed detections

  if (length( unlist(DetectionList)) > 0 &  length( unlist(DetectionList)) < nrow(TestDataTable) ) {

    missed_detections <- TestDataTable[-unlist(DetectionList), ]
    # Prepare missed detections data
    missed_detections <- missed_detections[, c("Selection", "View", "Channel", "Begin.Time..s.", "End.Time..s.", "Low.Freq..Hz.", "High.Freq..Hz.")]
    missed_detections <- missed_detections
    missed_detections$File.Name <- ShortName
    missed_detections$model.type <- 'SVM'
    missed_detections$probability <- 0
    missed_detections$signal <- 'female.gibbon'

    # Append missed detections to TopModelDetectionDF
    TopModelDetectionDF <- rbind.data.frame(TopModelDetectionDF, missed_detections)
  }

  if (length( unlist(DetectionList)) == 0) {

    missed_detections <- TestDataTable
    # Prepare missed detections data
    missed_detections <- missed_detections
    missed_detections$File.Name <- ShortName
    missed_detections$model.type <- 'SVM'
    missed_detections$probability <- 0
    missed_detections$signal <- 'female.gibbon'

    # Append missed detections to TopModelDetectionDF
    TopModelDetectionDF <- rbind.data.frame(TopModelDetectionDF, missed_detections)

  }

}
}

# Check the resulting structure
head(TopModelDetectionDF)

# Create table of signals
table(TopModelDetectionDF$signal)
```


## Part 3c. Calculate performance metrics
Now that we have matched annotations and detections we can calculate performance over a number of thresholds
```{r,message=FALSE, warning=FALSE}
# Convert signal column to a factor variable
TopModelDetectionDF$signal <- as.factor(TopModelDetectionDF$signal)

# Display unique values in the signal column
unique(TopModelDetectionDF$signal)

# Define a vector of confidence Thresholds
Thresholds <-seq(0.1,1,0.1)

# Create an empty data frame to store results
BestF1data.framefemale.gibbonBinary <- data.frame()

# Loop through each threshold value
for(a in 1:length(Thresholds)){

  # Filter the subset based on the confidence threshold
  TopModelDetectionDF_single <-TopModelDetectionDF

  TopModelDetectionDF_single$Predictedsignal <-
    ifelse(TopModelDetectionDF_single$probability  <=Thresholds[a], 'noise','female.gibbon')

  # Calculate confusion matrix using caret package
  caretConf <- caret::confusionMatrix(
    as.factor(TopModelDetectionDF_single$Predictedsignal),
    as.factor(TopModelDetectionDF_single$signal),positive = 'female.gibbon',
    mode = 'everything')


  # Extract F1 score, Precision, and Recall from the confusion matrix
  F1 <- caretConf$byClass[7]
  Precision <- caretConf$byClass[5]
  Recall <- caretConf$byClass[6]
  FP <- caretConf$table[1,2]
  TN <- sum(caretConf$table[2,])#+JahooAdj
  FPR <-  FP / (FP + TN)
  # Create a row for the result and add it to the BestF1data.frameGreyfemale.gibbon
  #TrainingData <- training_data_type
  TempF1Row <- cbind.data.frame(F1, Precision, Recall,FPR)
  TempF1Row$Thresholds <- Thresholds[a]
  BestF1data.framefemale.gibbonBinary <- rbind.data.frame(BestF1data.framefemale.gibbonBinary, TempF1Row)
}
```

## 3d. Plot the results
We can plot precision, recall, and F1 as a function of the model confidence or threshold. Note that we would not expect really great performance, as we only have a small number of training samples. 
```{r, warning=FALSE, message=FALSE }

female.gibbonMax <- round(max(na.omit(BestF1data.framefemale.gibbonBinary$F1)),2)

# Metric plot
female.gibbonBinaryPlot <- ggplot(data = BestF1data.framefemale.gibbonBinary, aes(x = Thresholds)) +
  geom_line(aes(y = F1, color = "F1", linetype = "F1")) +
  geom_line(aes(y = Precision, color = "Precision", linetype = "Precision")) +
  geom_line(aes(y = Recall, color = "Recall", linetype = "Recall")) +
  labs(title = paste("Female gibbon automated detection (binary) \n max F1:",female.gibbonMax),
       x = "Confidence",
       y = "Values") +
  scale_color_manual(values = c("F1" = "blue", "Precision" = "red", "Recall" = "green"),
                     labels = c("F1", "Precision", "Recall")) +
  scale_linetype_manual(values = c("F1" = "dashed", "Precision" = "dotted", "Recall" = "solid")) +
  theme_minimal() +
  theme(legend.title = element_blank())+
  labs(color  = "Guide name", linetype = "Guide name", shape = "Guide name")

female.gibbonBinaryPlot
```
