---
title: "gibbonR: An R package for the automated detection and classification of female gibbon calls from long-term acoustic recordings"
author: "Dena J. Clink"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalikebookd
link-citations: yes
github-repo: DenaJGibbon/gibbonR-tutorial
# output:
#   html_document:
#     toc: true
#     toc_depth: 2
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Package installation {-}
You can install the development version from [GitHub](https://github.com/DenaJGibbon) with:
```{r, eval=FALSE}
# install.packages("devtools")
# devtools::install_github("DenaJGibbon/gibbonR")

library(gibbonR)
```

# Package summary {-}
'gibbonR' has functions to:     
  + detect sound events using band-limited energy summation (DetectBLED)
  + detect and classify sound events using random forest or support vector machine algorithms (gibbonR)   
  + visualize sound events using UMAP plots with affinity propagation lcustering (gibbonID)
